---
title: "MTH6139 Time Series" 
subtitle: "Coursework 1" 
author: "Muhammad Sameed" 
date: "Spring term 2026" 
output: 
  html_document:
    toc: true
    toc_float: true
    theme: spacelab 
    highlight: tango
---
 
```{r, echo=FALSE}
# This code will display the QMUL logo at the top right of the page
# Do not change this code
htmltools::img(src = knitr::image_uri("images/QMlogo.png"),
               alt = 'logo',
               style = 'position:absolute; top:0; right:0; padding:10px; width:20%;')
```



# Section 1: Introduction

This project is about analysing the number of practical driving test passes in the UK from Jan 2019 to Nov 2024.


```{r, message=FALSE, warning=FALSE}
library(prophet)

raw_data = read.csv("car-practical-test-pass-rates-jan-2019-to-nov-2024.csv")


```

# Preparing the data 
Preparing the data for Prophet using the two columns {ds,y} = {Time , Passes}
```{r}
test_passes = data.frame(
    ds = as.Date(raw_data$Time, format = "%d/%m/%Y"),
    y = as.numeric(gsub(",", "", raw_data$Passed))
)

```
The "date" column is called ds and "passed" column is labelled y.

# Plotting the Time Series

```{r}
plot(test_passes$ds, test_passes$y,
     type = "l",
     main = "Number of Practical Driving Passes",
     xlab = "Date",
     ylab = "Passes",
     xaxt = "n")

axis.Date(1,
          at = seq(min(test_passes$ds),
                   max(test_passes$ds),
                   by = "1 year"),
          format = "%Y")
```        

From the Time Series we can see a big fall in number of passes around 2020. This could be related to the COVID period, where driving tests were disrupted. There is still a backlog on driving tests availability even after 6 years.However, after this the number of passes have graudally increased.
          
# Forecasting With Prophet
```{r}
model = prophet(test_passes)

future = make_future_dataframe(model,
                                periods = 12,
                                freq = "month")

forecast = predict(model, future)
plot(model, forecast)
```



This predicts the number of future practical driving test passes for the next year.

Looking at the graph, the number of test passes falls drastically around 2020, which may be due to COVID, however ever since then it has gradually grown.
